### 2021-10-14 paper/regularization.tex
changes

\documentclass[twoside]{article}

\usepackage{aistats2022}
% If your paper is accepted, change the options for the package
% aistats2022 as follows:
%
%\usepackage[accepted]{aistats2022}
%
% This option will print headings for the title of your paper and
% headings for the authors names, plus a copyright note at the end of
% the first column of the first page.

% If you set papersize explicitly, activate the following three lines:
%\special{papersize = 8.5in, 11in}
%\setlength{\pdfpageheight}{11in}
%\setlength{\pdfpagewidth}{8.5in}

% If you use natbib package, activate the following three lines:
%\usepackage[round]{natbib}
%\renewcommand{\bibname}{References}
%\renewcommand{\bibsection}{\subsubsection*{\bibname}}

% If you use BibTeX in apalike style, activate the following line:
%\bibliographystyle{apalike}
%\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{bm}
\usepackage[round]{natbib}
\usepackage[inline]{enumitem}
\usepackage{tikz}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{mwe}

\usepackage{array}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\newcolumntype{C}[1]{>{\PreserveBackslash\centering}p{#1}}
\newcolumntype{R}[1]{>{\PreserveBackslash\raggedleft}p{#1}}
\newcolumntype{L}[1]{>{\PreserveBackslash\raggedright}p{#1}}

@@

%\theoremstyle{definition}
\newtheorem{assumption}{Assumption}
\newtheorem{problem}{Problem}
\newtheorem{defn}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{fact}{Fact}

@@

\newcommand{\parent}[1]{\texttt{parent}({#1})}

\renewcommand{\star}[1]{{#1}^{*}}
\newcommand{\bad}[1]{{#1}^{\textit{bad}}}
\newcommand{\trans}[1]{{#1}^{T}}
\newcommand{\loss}{\ell}
\newcommand{\aaa}{\mathbf a}
\newcommand{\vv}{\mathbf v}
\newcommand{\uu}{\mathbf u}
\newcommand{\w}{\mathbf w}
\newcommand{\x}{\mathbf x}
\newcommand{\y}{\mathbf y}
\newcommand{\lone}[1]{{\lVert {#1} \rVert}_1}
\newcommand{\ltwo}[1]{{\lVert {#1} \rVert}_2}
\newcommand{\lp}[1]{{\lVert {#1} \rVert}_p}
\newcommand{\linf}[1]{{\lVert {#1} \rVert}_\infty}
\newcommand{\lF}[1]{{\lVert {#1} \rVert}_F}

@@

\newcommand{\ignore}[1]{}

@@

% If your paper is accepted and the title of your paper is very long,
% the style will print as headings an error message. Use the following
% command to supply a shorter title of your paper so that it can be
% used as headings.
%
%\runningtitle{I use this title instead because the last one was very long}

% If your paper is accepted and the number of authors is large, the
% style will print as headings an error message. Use the following
% command to supply a shorter version of the authors names so that
% they can be used as headings (for example, use only the surnames)
%
%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}

\twocolumn[

\aistatstitle{The Tree Loss: Improving Generalization with Many Classes}

\aistatsauthor{ Yujie Wang \And Michael Izbicki  }

\aistatsaddress{ Claremont Graduate University \And  Claremont McKenna College } ]

\begin{abstract}
    %Multiclass logistic regression has generalization error $O(\sqrt{k/n})$, where $k$ is the number of classes and $n$ the number of data points.

    Multiclass classification problems often have many semantically similar classes.
    For example, 90 of ImageNet's 1000 classes are for different breeds of dog.
    We should expect that these semantically similar classes will have similar parameter vectors,
    but the standard cross entropy loss does not enforce this constraint.
    %These classes are semantically similar to each other,
    %and so we expect their learned parameter vectors to also be similar.
    %Unfortunately, the standard cross entropy loss function does not enforce this constraint.
    %It has generalization error $O(\sqrt{k/n})$, where $k$ is the number of classes and $n$ the number of data points.

    We introduce the tree loss as a drop-in replacement for the cross entropy loss.
    The tree loss re-parameterizes the parameter matrix in order to guarantee that semantically similar classes will have similar parameter vectors.
    Using simple properties of stochastic gradient descent,
    we show that the tree loss's generalization error is asymptotically better than the cross entropy loss's.
    We then validate these theoretical results on synthetic data, image data (CIFAR100, ImageNet), and text data (Twitter).

    %Whereas The standard cross entropy loss function has generalization error $O(\sqrt{k/n})$,
    %where $k$ is the number of classes and $n$ the number of data points.

%Softmax Cross Entropy loss function treats every class equally while classes often have an internal structure. 
%We introduce cover tree loss function which takes advantage of this internal class structure.
%The cover tree loss works by reparameterizing the weight matrix for the cross entropy loss.
%Compare to cross entropy which has generalization error $O(\sqrt{k/n})$, the generalization error of cover tree loss is $O(\sqrt{1/n})$ which is independent of the number of classes. 
%We provide the theoretical justification and conduct extensive experiments on fully synthetic data and real world data. 
%For real world data, we take CIFAR100 dataset, IMAGENET dataset and TWITTER dataset.
%The cover tree loss outperforms than cross entropy loss function on metrics.
\end{abstract}

@@

When people do a multi-class classification task, the standard loss function is cross entropy.
The mechanism of cross entropy loss function is to maximize the assigned target class probability and punishes every misclassification equally, independent of some other information about the predicted class.
However, classes often have an internal structure. 
For example, dog and cat belong to same 'parent': animal and many transportations have same 'parent' while they have different brothers. Take the internal structure in the training procedure would allow the classifier to make less sever mistakes as it learns to predict similar classes.
In terms of theoretical result, the parameter matrix of cross entropy loss is bounded by $\sqrt{k}B$ (where $k$ is the number of classes) and for $n$ data points, the generalization error is $O(\sqrt{k/n})$ which is dependent on $k$.

In this paper, we introduce the tree loss which take advantage of the internal structure bewtween classes.
The idea of the tree loss is that we use tree structure to reduce the bound of the size of the parameter space and then the generalization error of the tree loss is $O(\sqrt{1/n})$ which is independent of $k$.
In the tree loss, we take a tree structure which defined by similarity between classes to re-parameterize the weight matrix for the cross entropy loss.
We incorporate class similarities with loss function to make the model learn more information about classes.
It can benefit the tasks on unbalanced datasets.
Take a emoji prediction task as an example, people often use emojis in tweets.
However, some emojis are popular and some emojis are rarely used.
This leads to an unbalanced tweets data and there is not enough training data for unpopular emojis.
The tree loss can learn more information and work well for unbalanced data.

Our contributions are as below:
\begin{itemize}
    \item [1] 
    We introduce the tree loss which take advantage of internal structure between all classes.
    \item [2] 
    We provide theoretical justification that the generalization error of the tree loss is independent of the number of classes.
    \item [3]
    We evaluate our tree loss on experiments. One is experiment on fully synthetic data which is designed to be convex. The another one is on real world data: CIFAR100, IMAGENET and TWITTER dataset.

\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\resizebox{\columnwidth}{!}{
\begin{tikzpicture}
    [ level distance=1.5cm
    , level 1/.style={sibling distance=3cm}
    , level 2/.style={sibling distance=1.5cm}
    ]
\node[draw, rounded corners=0.1in, inner sep=0.1in] at (-1.5in,0){\textbf{$U$-Tree}};
\node {\texttt{truck}}
    child {node {\texttt{boxer}}
      child {node {\texttt{tiger}}}
      child {node {\texttt{skunk}}}
      child {node {\texttt{boxer}}
        child {edge from parent[draw=none]} % Added
        child {node {\texttt{bulldog}}}
        child {node {\texttt{boxer}}}
        child {node {\texttt{husky}}}
        child {node {\texttt{sheepdog}}}
      }
      child {node {\texttt{bear}}}
      child {edge from parent[draw=none]} % Added
      }
    child {node {\texttt{truck}}
      child {edge from parent[draw=none]} % Added
      child {node {\texttt{truck}}}
      child {node {\texttt{bus}}}
    }
    child {node {\texttt{toaster}}}
    ;

\end{tikzpicture}
}
~~~
\resizebox{\columnwidth}{!}{
\begin{tikzpicture}
    [ level distance=1.5cm
    , level 1/.style={sibling distance=3cm}
    , level 2/.style={sibling distance=1.5cm}
    ]
\node[draw, rounded corners=0.1in, inner sep=0.1in] at (-1.5in,0){\textbf{$V$-Tree}};
\node {\textit{pseudo1}}
    child {node {\textit{pseudo2}}
      child {node {\texttt{tiger}}}
      child {node {\texttt{skunk}}}
      child {node {\textit{pseudo3}}
        child {edge from parent[draw=none]} % Added
        child {node {\texttt{bulldog}}}
        child {node {\texttt{boxer}}}
        child {node {\texttt{husky}}}
        child {node {\texttt{sheepdog}}}
      }
      child {node {\texttt{bear}}}
      child {edge from parent[draw=none]} % Added
      }
    child {node {\textit{pseudo4}}
      child {edge from parent[draw=none]} % Added
      child {node {\texttt{truck}}}
      child {node {\texttt{bus}}}
    }
    child {node {\texttt{toaster}}}
    ;

\end{tikzpicture}
}
    \caption{
        Example label tree structures for a subset of 10 ImageNet classes.
        The $U$-tree has class labels that repeat at multiple levels,
        and the $V$-tree introduces ``pseudoclasses''.
        The pseudoclass \textit{pseudo3} represents the class of ``dogs'',
        and the pseudoclass \textit{pseudo2} represents the class of ``animals''.
    }
    \label{fig:labeltree}
\end{figure*}

\section{The Tree Loss}

This section first formally defines our problem setting,
then formally defines the tree loss.
We conclude with a comparison between the tree loss and related loss functions.

\subsection{Problem Setting}

We consider the multi-class classification setting with $k$ classes and $d$ input feature dimensions.
The cross entropy loss is the standard loss function for this setting.
It is defined to be
\begin{equation}
    \label{eq:xentropy}
    \ell(W;(\x,y)) = - \log \frac {\exp(-\trans\w_y \x)}{\sum_{j=1}^k \exp(-\trans \w_j \x)}
\end{equation}
where for each class $i\in[k]$,
$\w_i : \R^d$ is the parameter vector associated with class $i$;
the variable $W : \R^{k \times d} = (\w_1; \w_2; ...; \w_k)$ is the full parameter matrix;
$\x : \R^d$ is the input feature vector;
and $y \in [k]$ is the input class label.

\subsection{The Tree Loss}

The tree loss reparameterizes the weight matrix used in the cross entropy loss in order to enforce that semantically similar labels will have similar weight vectors.
The reparameterization is defined according to a \emph{label tree} structure that captures the similarity between labels.
And we propose two variants of the tree loss called the $U$ and $V$ variants that use slightly different label tree structures.
The main purpose of the $U$-tree structure is to simplify the analysis and exposition of the $V$-tree structure,
and we show later in our theoretical results that the $V$-tree is strictly better than the $U$-tree.
Figure \ref{fig:labeltree} shows examples of both.

\subsubsection{The $U$-Tree Loss}

In a $U$-tree, every node is a class and the class of all internal nodes matches the class of one of its children.
For each class $i$, we define $\parent{i}$ to be the parent of the upper-most node for class $i$,
and the $\texttt{parent}$ of the root is the 0 vector..
For example, in the example of Figure \ref{fig:labeltree},
we have
\begin{equation*}
\begin{split}
    \parent{\texttt{husky}} &= \texttt{boxer} \text{, and}\\
    \parent{\texttt{bear}} &= \texttt{boxer} \text{, and}\\
    \parent{\texttt{boxer}} &= \texttt{truck} .\\
\end{split}
\end{equation*}
We associate for each class $i$ a new parameter vector $\uu_i$ and set
\begin{equation}
    \w_i = \w_{\parent{i}} - \uu_i,
    \label{eq:uu}
\end{equation}
and we define the parameter matrix $U : \mathbb R^{k\times d}$ to be $(\uu_1;...;\uu_k)$.
The $U$-tree loss is then defined by substituting \eqref{eq:uu} into \eqref{eq:xentropy} to get
\begin{equation*}
    \ell(U;(\x,y)) = - \log \frac {\exp(-(\trans\uu_{\parent{y}}-\trans\uu_y) \x)}{\sum_{j=1}^k \exp(-(\trans\uu_{\parent{j}}-\trans \uu_j) \x)}
\end{equation*}
We use the same $\ell$ notation for the standard cross entropy loss function and the $U$-Tree loss because the function is exactly the same;
the only difference is how we represent the parameters.

\subsubsection{The $V$-Tree Loss}

The $V$-tree is constructed from the $U$-tree by replacing all non-leaf nodes with ``pseudoclasses''.
Let $k'$ be the total number of nodes in the $V$-tree;
so $k'$ equals $k$ plus the number of internal nodes.
For each class $i$, we let the sequence $P_i$ denote the path from the leaf node to the root.
For example, using the $V$-tree in Figure \ref{fig:labeltree},
we have
\begin{equation*}
\begin{split}
    P_{\texttt{sheepdog}} &= (\texttt{sheepdog}, \textit{pseudo3}, \textit{pseudo2}, \textit{pseudo1} ) \\
    P_{\texttt{truck}} &= (\texttt{truck}, \textit{pseudo4}, \textit{pseudo1} )
\end{split}
\end{equation*}
Next, we associate a vector $\vv_i$ to each node $i$ in the tree structure.
In particular, there is a vector $\vv_i$ for each class and for each pseudo-class,
so there are more parameter vectors for the $V$-tree loss than either the $U$-tree loss or the standard cross entropy loss.

Finally, we reparameterize the weights for the cross entropy loss as the sum of the vectors along the paths.
Specifically, we set the weight vector for class $i$ to be 
\begin{equation}
    \label{eq:covertree-reparam}
    \w_j = \sum_{j\in P_i} \vv_i
    .
\end{equation}
The $V$-tree loss is defined by substituting \eqref{eq:covertree-reparam} into \eqref{eq:xentropy} to get
\begin{equation}
    \ell(V;(\x,y)) = - \log \frac {\exp(-\sum_{k\in P_y}\trans\vv_k \x)}{\sum_{j=1}^k \exp(- \sum_{k\in P_j}\trans\vv_k \x)}
    .
\end{equation}
We continue to use the $\ell$ function to represent both the standard cross entropy loss and the $V$-tree loss function to emphasize that these are the same loss functions,
just with different parameterizations of the weight matrix.

\subsection{Intuition}

The intuition behind the $V$-tree reparameterization is that whenever we perform a gradient update on a data point with class $i$,
we will also be ``automatically'' updating the parameter vectors of similar classes.
To see this, note that when we have a \texttt{sheepdog} data point,
we will perform gradient updates on all $\vv_i$ in $P_\texttt{sheepdog}$; i.e. $\vv_{sheepdog}$, $\vv_{\textit{pseudo3}}$, $\vv_{\textit{pseudo2}}$, and $\vv_{\textit{pseudo1}}$.
This will cause the $\w_{\texttt{husky}}$ and $\w_{\texttt{bear}}$ parameters (among many others) to update because they also depend on the pseudoclass vectors $\vv_{\textit{pseudo3}}$ and $\vv{_\textit{pseudo2}}$.
Because $P_\texttt{husky}$ has a larger overlap with $P_\texttt{sheepdog}$ than $P_\texttt{bear}$,
the parameters of this more similar class will be updated more heavily.

\subsection{Implementation Notes}

Both the $U$-tree and $V$-tree losses are easy to implement in practice using the built-in cross entropy loss function in libraries like PyTorch \citep{NEURIPS2019_9015} or Tensorflow \citep{tensorflow2015-whitepaper}.
The only change needed is to define the loss in terms of the sum of vectors $V$ instead of a variable tensor $W$.
This ensures that the loss is implemented efficiently and is numerically stable.

In practice, the $V$-tree loss is slightly slower than the standard cross entropy loss and $U$-tree losses due to the additional summation over the paths.
In our experiments, we found this this to cause the $V$-tree loss to be about twice as slow.

\section{Theoretical Results}

It is well known that when using the standard cross entropy loss,
SGD converges at a rate of $O(\sqrt{k/n})$.
Our theoretical results asymptotically improve this convergence rate.

With an appropriately constructed label tree,
we show convergence at a rate of $O(\sqrt{\log k/n})$ or $O(\sqrt{k^{1-1/c}/n})$,
where $c$ is a measure of difficulty of the problem.
But what happens if the tree is poorly constructed?
Remarkably, nothing bad.
We also show that for arbitrarily bad label tree structures,
the $U$ and $V$ losses maintain a $O(\sqrt{k/n})$ convergence rate.
This suggests that there is essentially no downside to using the tree loss parameterizations.

\fixme{
Our analysis relies on simple properties of stochastic gradient descent,
and we view the simplicity of our analysis as a major strength of the tree loss.
In this section, we first review standard results of stochastic gradient descent needed for our analysis.
Then we formally state and prove our main result.
}

\subsection{SGD Background}

SGD is a standard algorithm for optimizing convex losses,
and it has many variations.
Our analysis uses the results from Chapter 14 of \citet{shalev2014understanding} which are based off of a version of SGD that uses parameter averaging.\footnote{
Other variants of SGD have similar results,
and our results are easy to generalize to other SGD variants.
We refer the reader to \citet{shalev2014understanding} for a discussion of these variants.}
In this section we recall these results,
occasionally changing the names of variables to avoid conflicts with variable names used elsewhere in this paper.

We assume we are optimizing a function $f : \Theta \to \mathbb R$,
and that we are only given noisy samples of the function and its gradient.
We let $\theta^{t}$ denote the parameter value at iteration $t$,
and $\eta : \mathbb R$ be the step size.
Then the parameter update for step $t$ is defined to be
\begin{equation}
    \theta^{(t+1)} = \theta^{(t)} - \eta \nabla f(\theta^{(t)}).
\end{equation}
When using parameter averaging, the final model parameters are defined to be
\begin{equation}
    \bar\theta = \frac1T \sum_{i=1}^T \theta^{(t)}
    .
\end{equation}
Parameter averaging is not often used in practice because of the memory requirements,
but we use it here due because theorems using parameter averaging are typically much easier to state and prove.
Equivalent results without parameter averaging typically require and additional $\log T$ factor in their convergence rates\citep{}.

The main result we rely on is a bound on the convergence rate of SGD for convex Lipschitz functions.
We reproduce Theorem 14.8 of \citet{shalev2014understanding} in the theorem below.

\begin{theorem}
    \label{thm:sgd}
    Let $\beta,\rho>0$.
    Let $f$ be a convex function and let $\star \theta = \argmin_{\theta : \ltwo{\theta}\le \beta} f(\theta)$.
    Assume that SGD is run for $T$ iterations with $\eta = \sqrt{\frac{\beta^2}{\rho^2 T}}$.
    Assume that for all $t$, $\ltwo{\nabla f(\theta^{(t)})} \le \rho$.
    Then,
    \begin{equation}
        \E f(\bar \theta) - f(\star \theta) \le \frac{\beta\rho}{\sqrt{T}}
        .
    \end{equation}
\end{theorem}

\subsection{SGD and Statistical Learning}

We now introduce more statistical learning theory notation,
then apply Theorem \ref{thm:sgd} above to bound the rate of learning.

Define the true loss of our learning problem as
\begin{equation}
    L_D(A) = \E_{(\x,y)\sim D} \ell(A; (\x, y))
\end{equation}
where $A \in \{W, U, V\}$ is a parameterization of the cross entropy loss.
The optimal parameter vector is defined to be the minimum of the true loss:
\begin{equation}
    \star A = \argmin_{A} L_D(A).
\end{equation}
In order to apply Theorem \ref{thm:sgd} to a learning problem,
we make the following assumption about our the size of our data points:
\begin{assumption}
    \label{ass:lip}
    For each feature vector $\x$ in the data set, 
        $\ltwo{\x} \le \rho$
        .
\end{assumption}
Under this assumption, the cross entropy loss $\ell$ is $\rho$-Lipschitz for all parameterizations,
and the true loss $L_D$ is therefore also $\rho$-Lipschitz.

Applying Theorem \ref{thm:sgd} with $f=L_D$ gives
\begin{corollary}
    \label{ref:cor:A}
    Under Assumption \ref{ass:lip}, we have that for any parameterization $A$ of the cross entropy loss,
    \begin{equation}
        \E L_D(\bar A) - L_D(\star A) \le \frac{\lF{\star A}\rho}{\sqrt n}.
        \label{eq:Aconv}
    \end{equation}
\end{corollary}

We can now bound the convergence of the standard cross entropy loss and the $U$ and $V$-tree losses by bounding the Frobenius norms of their parameter matrices.

In order to recover the standard convergence rate of multiclass classification, we make the following assumption.
\begin{assumption}
    \label{ass:B}
    For each class $i$, the optimal parameter vector $\star\w_i$ satisfies
        $\ltwo{\star\w_i} \le B$
        .
\end{assumption}
Note that this assumption is not limiting because the value of $B$ need not be known during the optimization procedure.

It follows that
\begin{equation}
    \lF{\star W}^2 = \sum_{i=1}^k \ltwo{\star\w_i} \le kB^2.
    \label{eq:starW}
\end{equation}
Substituting Eq \eqref{eq:starW} into \eqref{eq:Aconv} gives the following bound.

\begin{theorem}
\label{theorem:xentropy}
    Under assumptions \ref{ass:lip} and \ref{ass:B} above,
then the parameter vector $\bar W$ estimated from one epoch of SGD satisfies
\begin{equation}
    \E L_D(\bar W) - L_D(W^*)
    \le \frac {\sqrt kB\rho}{\sqrt n}
    .
\end{equation}
\end{theorem}
\fixme{
It is important to note that this result above is an upper bound and not a lower bound.
This $O(\sqrt{k/n})$ convergence rate matches the convergence rate implied by the Nataranj dimension \citep{}.
We are not aware of any lower bounds for SGD in the Lipschitz setting.
\citet{nguyen2018tight,jentzen2020lower} study the strongly convex case and find that the upper bounds match the lower bounds.
}

\subsection {Bad Tree Structures}

\begin{lemma}
    \ref{lemma:starV}
    We have that for all $V$-trees derived from a $U$-tree,
    \begin{equation}
        \lF{\star V} \le \lF{\star U}.
    \end{equation}
\end{lemma}
\begin{proof}
    \fixme{}
\end{proof}

The following lemma shows that even for arbitrarily bad tree structures,
the bound on the parameter spaces cannot be too bad.

\begin{lemma}
    \ref{lemma:starU}
    Under assumption \ref{ass:B},
    the following bound holds for all $U$-tree structures:
    \begin{equation}
        \lF{\star U} \le 2\sqrt{k}B.
    \end{equation}
\end{lemma}
\begin{proof}
    We have
    \begin{align}
        \lF{\star U}^2 
        &= \sum_{i=1}^k \ltwo{\star\uu_i}^2 \\
        &= \sum_{i=1}^k \ltwo{\star\w_i - \star\w_{\parent{i}}}^2 \\
        &\le \sum_{i=1}^k \left(\ltwo{\star\w_i} + \ltwo{\star\w_{\parent{i}}} \right)^2 \\
        &\le \sum_{i=1}^k (2B)^2 \\
        &\le k (2B)^2
        .
    \end{align}
\end{proof}

\subsection{Good Tree Structures}

\begin{assumption}
We are given a distance metric $d$ over the labels such that for all labels $i$ and $j$,
\begin{equation}
    \frac 1 \lambda d(i,j)
    \le \ltwo{\star \w_i - \star \w_j}
    \le \lambda d(i, j).
\end{equation}
We denote by $c$ the doubling dimension of the metric $d$.
\end{assumption}

\begin{lemma}
    \ref{lemma:main}
    When $c\le1$, we have that
    \begin{equation}
        \lF{\star U} \le \tfrac{1}{\sqrt2}\lambda B \sqrt{\log_2 k}.
        \label{eq:c<=1}
    \end{equation}
    When $c>1$, we have that
    \begin{equation}
        \lF{\star U} \le \sqrt{5}\lambda B \sqrt{k^{(1-1/c)}}.
        \label{eq:c>1}
    \end{equation}
\end{lemma}
\begin{proof}
%Construct a cover tree over the distance metric $d$.
%For each class $i$, define the vector
%\begin{equation}
    %\uu_i =
    %\begin{cases}
        %\w_i & \text{if $i$ is the root of the cover tree} \\
        %\w_i - \w_{\parent{i}} & \text{otherwise} \\
    %\end{cases}
    %.
%\end{equation}
%Define the vector $\star\uu_i$ similarly by replacing all the $\w_i$ with $\star\w_i$ above.
    First, we review some simple properties of the cover tree structure that will be needed in our proof.
    Let $q$ denote the maximum number of children of any node in the cover tree.
    A standard property of cover trees is that $q \le (4^c)$ \citep[Lemma 25]{izbickithesis}.
    Let $Q_i$ denote the number of nodes at level $i$ in the cover tree.
    Then we have that $|Q_i| \le q^i.$
    Finally, let $r$ be the height of the cover tree.
    Without loss of generality, we analyze the bound when the cover tree is perfectly balanced and so $r = \log_q k = \log_2 k / (2c)$.

We have that
\begin{align}
    \lF{\star U}^2
    &= \sum_{i=1}^k \ltwo{\star\uu_i}^2 \\
    &= \sum_{i=0}^r \sum_{j\in Q_i} \ltwo{\star\uu_j}^2 \\
    &= \sum_{i=0}^r \sum_{j\in Q_i} \ltwo{\star\w_j - \star\w_{\parent{j}}}^2 \\
    &\le \sum_{i=0}^r \sum_{j\in Q_i} \lambda^2 d(j, \parent{j})^2 \\
    &\le \sum_{i=0}^r \sum_{j\in Q_i} \lambda^2 (B/2^i)^2 \\
    &\le \sum_{i=0}^r q^i \lambda^2 (B/2^i)^2 \\
    &\le \sum_{i=0}^r \lambda^2 B^2 (q/4)^i 
    \label{eq:precases}
\end{align}
We proceed by considering the case of $c\le1$ and $c>1$ separately.

When $c\le1$, $q\le4$.
Substituting into Eq \eqref{eq:precases}, we have
\begin{align}
    \lF{\star U}^2
    &\le \sum_{i=0}^r \lambda^2 B^2 \\
    &= (r+1) \lambda^2 B^2 \\
    &= \frac{\log_2k}{2c} \lambda^2 B^2 \\
    &\le \frac{\log_2k}{2} \lambda^2 B^2
    .
\end{align}
This matches Eq \eqref{eq:c<=1} in the lemma's statement.

When $c>$, then $q>4$.
We first show the following equality.
\begin{align}
\left(\frac q 4\right)^r
&=
\left(\frac q 4\right)^{\log_q k}\\
&=
\frac{k}{4^{\log_q k}}\\
&=
\frac{k}{4^{{\log_4 k}/{\log_4 q}}}\\
&=
\frac{k}{k^{1/\log_4 q}}\\
&=
k^{\left(1 - {1}/{\log_4 q}\right)} \\
&=
k^{\left(1 - 1/c\right)}
.
\label{eq:lFU:frac}
\end{align}
Substituting into Eq \eqref{eq:precases}, we have
\begin{align}
    \lF{\star U}^2
    &= \lambda^2B^2 \left(\frac{(q/4)^{r+1}-1}{q/4-1}\right) \\
    \label{eq:lFU:4}
    &= \lambda^2B^2 \left(\frac{(q/4)k^{(1-1/c)}-1}{q/4-1}\right) \\
    &\le \lambda^2B^2 \left(\frac{(q/4)k^{(1-1/c)}}{q/4-1}\right) \\
    &\le 5 \lambda^2B^2 k^{(1-1/c)}
    \label{eq:lFU:5}
\end{align}
    Inequality \eqref{eq:lFU:5} follows because $\tfrac{q/4}{q/4-1} \le 5$ for integer values of $q>4$.
\end{proof}

%Substituting \eqref{eq:lFU:frac} info \eqref{eq:lFU:5} gives
%\begin{align}
    %\lF{U} \le \lambda B \sqrt{\frac{\log_2 k}{2c} + 1}\sqrt{k^{(1-1/c)}}
%\end{align}
%\begin{lemma}
    %\label{lemma:technical}
    %For all $a>1$ and $b>1$, we have that
    %\begin{equation}
        %\frac {1-a^b}{1-a} \le ba^b
    %\end{equation}
%\end{lemma}

\begin{figure*}
\includegraphics[width=\columnwidth,height=1.5in]{fig/images/accuracy_vs_n.png}
\includegraphics[width=\columnwidth,height=1.5in]{fig/images/accuracy_vs_d.png}
\includegraphics[width=\columnwidth,height=1.5in]{fig/images/accuracy_vs_class.png}
\includegraphics[width=\columnwidth,height=1.5in]{fig/images/accuracy_vs_sigma.png}
\caption{
    Results of our Synthetic Experiments I.
    The x-axis of each plot shows which problem parameter is being varied.
    As our theory predicts, the Tree Loss outperforms the baseline loss functions in all data regimes.
}
\label{fig:synth:1}
\end{figure*}

\section{Experiments}

\subsection{Synthetic data}

In this experiment we will demonstrate the tree loss's effectiveness on a simple multiclass logistic regression problem.
The problem will be specifically designed to be convex and have low intrinsic dimension in the class labels in order to illustrate the advantages of the cover tree loss.

\subsubsection{Data Generation Procedure}
\label{sec:exp:synth:problem}

Let $\mathcal N$ be the standard normal distribution.
Then sample the true parameter matrix
\begin{align}
    \star W \sim \mathcal N^{k\times d}
    .
\end{align}
For all $i \in [k]$, the row $\w_i^*$ of $W^*$ corresponds to the parameter vector for class $i$ and has dimension $d$.

For each data point $i\in[n]$,
we sample the data point according to the following rules:
\begin{align}
    y_i &\sim \text{Uniform}([k]), \text{and} \\
    \x_{i} &\sim \mathcal N(\w^*_{y_i}; \sigma),
\end{align}
where $\sigma$ controls the noise in the dataset.
Increasing $\sigma$ adds noise and makes classes harder to distinguish,
whereas decreasing $\sigma$ reduces noise and makes the classes easier to distinguish.

With our problem defined in this way, it is clear that the empirical risk minimizer
\begin{equation}
    \hat W = \argmin_{W} \tfrac 1 n \ell(W; \x_i, y_i)
\end{equation}
converges to $W^*$.

\subsubsection{Experiment I: Data Regimes}

In this experiment, we investigate the tree loss's performance in a wide range of settings as we vary the number of data points ($n$), the number of dimensions ($d$), number of class labels ($k$), and the Bayes error ($\sigma$) of our problem.
First, we construct a standard experimental setting with $n=100$, $d=64$, $k=10$, and $\sigma=1$.
Then, for each of these parameters, we investigate the effect of that parameter's performance on the problem by varying it over a large range.
For each value in the range, we:
(1) randomly sample 50 problems using the procedure described in Section \ref{sec:exp:synth:problem};
(2) for each problem, we:
(2a) sample a training set with $n$ data points and a test set with 10000 data points;
(2b) train 4 models on the training set using the standard cross entropy loss, tree loss, simloss, and hierarchical softmax;
(2c) report the average accuracy accross the sampled test sets.

Figure \ref{fig:synth:1} shows the results of these experiments.
As our theory suggests, the tree loss outperforms all other losses in all regimes.
Consider the top-left plot.
As the number of data points $n$ gets large,
then all of the models have accuracies that converge to 1;
for most problems, however, the number of data points will not be sufficiently large to drive the generalization error to 0,
and the fewer data points that the problem has,
the better the tree loss performs compared to other losses.

\fixme{
From figure \ref{fig.1}, it is clear that the tree loss achieves the best performance. 
The tree loss converges at a faster rate as number of data points and dimension of parameter matrix increase.
If we increase the number of classes, the tree loss can keep a higher accuracy and the accuracy of SimLoss is higher than cross entropy loss and hierarchical softmax.
It is as our expected, the tree loss can improve performance on large classes tasks.
Besides, when we increase the noise in the dataset, the accuracy of all loss functions decrease, yet the tree loss decreases slower.
Among the four loss functions, hierarchical softmax hurts the performance obviously. 
We would not apply hierarchical softmax on real world data experiment in terms of computation.
}

\subsubsection{Experiment II: Parameter Norms}

>>>>

\documentclass[10pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{natbib}
\usepackage[inline]{enumitem}
\usepackage{tikz}
\usepackage{booktabs}
\usepackage{subfigure}

@@

\theoremstyle{definition}
\newtheorem{problem}{Problem}
\newtheorem{defn}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{fact}{Fact}

@@

\newcommand{\trans}[1]{{#1}^{T}}
\newcommand{\loss}{\ell}
\newcommand{\Loss}{\mathcal{L}}
\newcommand{\aaa}{\mathbf a}
\newcommand{\vv}{\mathbf v}
\newcommand{\uu}{\mathbf u}
\newcommand{\w}{\mathbf w}
\newcommand{\x}{\mathbf x}
\newcommand{\y}{\mathbf y}
\newcommand{\lone}[1]{{\lVert {#1} \rVert}_1}
\newcommand{\ltwo}[1]{{\lVert {#1} \rVert}_2}
\newcommand{\lp}[1]{{\lVert {#1} \rVert}_p}
\newcommand{\linf}[1]{{\lVert {#1} \rVert}_\infty}
\newcommand{\lF}[1]{{\lVert {#1} \rVert}_F}

@@

@@

\begin{center}
\Huge
Tree Regularization
\end{center}

@@

This note introduces \emph{tree regularization}.
For multi-class learning problems, tree regularization achieves a generalization error of $O(\sqrt{\log k / m})$ which is comparable to the tree loss's error of $O(\sqrt{1/m})$.
Tree regularization has several advantages over the tree loss, however:
\begin{enumerate}
    \item
        It can work with any loss function.
        For example, we can apply tree regularization to multiple regression, multi-label prediction, and objection detection problems.
    \item
        The optimal tree structure can be learned at training time and the problem remains convex.
    \item
        The analysis works for all optimization algorithms, not just stochastic gradient descent.
\end{enumerate}

\section{Notation}
Many regularized learning problems take the form
\begin{equation}
    \argmin_{W} \Loss(Z ; W) + \lambda R(W)^2
\end{equation}
where
$Z = \{z_1, ..., z_m\}$ is the training set,
$W$ is the model parameters,
\begin{equation}
    \Loss(Z ; W) = \frac 1m \sum_{i=1}^m \loss(z_i; W)
\end{equation}
and $R : W \to \R^+$ is a regularization function.

We assume that the parameters $W$ are given in matrix as a $k \times d$ matrix.
In the multi-class prediction problem,
each row $i$ is a $d$ dimensional vector that represents the parameter matrix for class $i$.

\section{Tree Regularization}

The Tree Regularizer is defined to be
\begin{equation}
    R(W, V, A) 
    = \sum_{i} \left\lVert \w_i - \sum_{j}A_{i,j} \vv_j \right\lVert
    = \sum_{i} \left\lVert \w_i - A_i V \right\lVert
\end{equation}

The idea of the convergence proof is to show that any loss function that is $\rho$-Lipschitz with respect to the L2 norm is $\rho/k$ Lipschitz with respect to the tree regularizer (for fixed $V$,$A$ and near the optimal solution).
Then applying the exercise at the end of the chapter solves the problem.
